A Survey on Ordered Binary Decision Diagrams

### 2015-09-14 paper.tex
Add sifting algorithm example; Revise dynamic variable ordering description

%%% Packages
\usepackage{graphicx}
\usepackage{balance}
\usepackage{tikz}
\usepackage{standalone}
\usepackage{subfigure}

@@

@@

The sifting algorithm therefore uses a brute-force enumeration strategy. It starts
by swapping the processed variable $x$ downwards one level at a time, i.e. swapping 
it with its successor variable, which is located on the layer below in the OBDD.
This is done until $x$ is the next to last variable in the OBDD. Then, this is
done in the other way round, i.e. swapping $x$ with its predecessor variable until
$x$ becomes the top variable of the OBDD. While proceeding, the ideal position
is memorized and after $x$ became the top variable, $x$ is moved to this ideal
position by, again, swapping levels one by one.

@@

\begin{figure}[ht]
    \centering
    \includestandalone[height = 3cm]{figs/min-obdd}
    \label{fig:min-obdd}
    \caption{Minimized OBDD (optimal).}
\end{figure}
\tbr

>>>>

%%% Packages
\usepackage{graphicx}
\usepackage{balance}
\usepackage{tikz}
\usepackage{standalone}
\usepackage{subfigure}
\usepackage{colortbl}

@@

\newcommand\diag[4]{%
  \multicolumn{1}{p{#2}|}{\hskip-\tabcolsep
  $\vcenter{\begin{tikzpicture}[baseline=0,anchor=south west,inner sep=#1]
  \path[use as bounding box] (0,0) rectangle (#2+2\tabcolsep,\baselineskip);
  \node[minimum width={#2+2\tabcolsep-\pgflinewidth},
        minimum  height=\baselineskip+\extrarowheight-\pgflinewidth] (box) {};
  \draw[line cap=round] (box.north west) -- (box.south east);
  \node[anchor=south west] at (box.south west) {#3};
  \node[anchor=north east] at (box.north east) {#4};
 \end{tikzpicture}}$\hskip-\tabcolsep}}

@@

The sifting algorithm therefore uses a brute-force enumeration strategy to find
the optimal variable position. It starts by swapping the processed variable $x$
downwards one level at a time, i.e. swapping it with its successor variable, which
is located on the layer below in the OBDD. This is done until $x$ is the next to
last variable in the OBDD. Then, this is done in the other way round, i.e. swapping
$x$ with its predecessor variable until $x$ becomes the top variable of the OBDD.
While proceeding, the ideal position is memorized and after $x$ became the top
variable, $x$ is moved to this ideal position by, again, swapping levels one by
one~\cite{RUDELL93}. Figure~\ref{fig:sifting-algorithm} shows this procedure
applied to an example, requiring a total of 9 steps for 5 possible positions.

\begin{figure}
    \centering
    \begin{tabular}{l|l|l|l|l|l|}
        \hline
        \diag{0.1em}{1cm}{\scriptsize{Step}}{\scriptsize{Pos}} & $1$ & $2$ & $3$ & $4$ & $5$ \tabularnewline
        \hline
        init & $v$ & $w$ & $x$ & $y$ & $z$ \tabularnewline
        \hline
        1 & $v$ & $w$ & \cellcolor{red!50}$y$ & \cellcolor{red!50}$x$ & $z$ \tabularnewline
        \hline
        2 & $v$ & $w$ & $y$ & \cellcolor{red!50}$z$ & \cellcolor{red!50}$x$ \tabularnewline
        \hline
        3 & $v$ & $w$ & $y$ & \cellcolor{red!50}$x$ & \cellcolor{red!50}$z$ \tabularnewline
        \hline
        4 & $v$ & $w$ & \cellcolor{red!50}$x$ & \cellcolor{red!50}$y$ & $z$ \tabularnewline
        \hline
        5 & $v$ & \cellcolor{red!50}$x$ & \cellcolor{red!50}$w$ & $y$ & $z$ \tabularnewline
        \hline
        6 & \cellcolor{red!50}$x$ & \cellcolor{red!50}$v$ & $w$ & $y$ & $z$ \tabularnewline
        \hline\hline
        7 & \cellcolor{green!50}$v$ & \cellcolor{green!50}$x$ & $w$ & $y$ & $z$ \tabularnewline
        \hline
        8 & $v$ & \cellcolor{green!50}$w$ & \cellcolor{green!50}$x$ & $y$ & $z$ \tabularnewline
        \hline
        9 & $v$ & $w$ & \cellcolor{green!50}$y$ & \cellcolor{green!50}$x$ & $z$ \tabularnewline
        \hline
    \end{tabular}
    \caption{Sifting algorithm (ideal position of $x$: 4)}
    \label{fig:sifting-algorithm}
\end{figure}

This procedure could be refined to choose the direction before actually start
swapping the processed variable. As a result, the variable is always swapped
towards the closer variable position first.

To obtain a proper variable ordering, variables are sorted descending in the number
of nodes per OBDD level. Then, the ideal position is computed for each variable
utilizing the approach described above. The sifting algorithm has one big advantage
over other similar algorithms, like \textit{window permutation algorithm}, namely
variables can move many levels because all positions are evaluated, regardless of
the temporary size of the OBDD.

@@

\begin{figure}[ht]
    \centering
    \includestandalone[height = 3cm]{figs/min-obdd}
    \caption{Minimized OBDD (optimal).}
    \label{fig:min-obdd}
\end{figure}
\tbr
